---
title: "A Predictive Model for ODI World Cup Matches"
author: "Krishna Kumar, Srihari Srinivasan"
date: "APR 30, 2023"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
library(tidyverse)
library(knitr)
library(tinytex)
library(cricketdata)
library(leaps)

knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(out.width = "60%")
```

## Introduction
Given that this is a One Day International (ODI) World Cup year, our analysis of batting and bowling performance metrics on the margin of victory, `mov`, is focused on prior World Cup matches. With knowledge of how the ODI format has changed since the introduction of the shorter, Twenty20 (T20) format, particularly the 2007 T20 World Cup, we have limited our data to the last five ODI World Cups (2003, 2007, 2011, 2015, and 2019). After wrangling ten batting metrics and four bowling metrics from 4774 individual player performances, we are looking to build a predictive model and gain insight as to which metrics, if any, have an effect on `mov`. In doing so, while this may be beyond the scope of this project, we hope to ultimately use our model to predict the outcome of the 2023 ODI World Cup.


## Exploratory Analysis

```{r, echo=FALSE}
odi <- read.csv("~/Downloads/odi.csv")
odi$date <- as.Date(odi$date)
odi$date <- format(odi$date, "%Y")
odi <- rename(odi, year = date)
```


## Model Development


## Model Analysis


## Conclusion

